---
title: "Polynomials tsne"
author: "Andrew Lampinen"
date: "March 11, 2020"
output: html_document
---

```{r}
library(tidyverse)
library(ggpubr)
library(tsne)
```
```{r}
theme_set(theme_classic())
```

# data loading

```{r}
results_dir = "./polynomial_rep_results"
num_runs = 5
```


```{r message=FALSE, warning=FALSE}
d = replicate(num_runs, data.frame())
for (run_i in 0:(num_runs-1)) {
  this_d = read_csv(sprintf("%s/run%i_task_representations.csv", results_dir, run_i)) 
  this_d = this_d %>%
    mutate(run = run_i)# %>%
    # gather(task, value, -run, -dimension) %>%
    # separate(task, c("task", "base_or_meta"), sep=":") %>%
    # mutate(base_or_meta = gsub("_[0-9]+", "", base_or_meta)) # some tasks are duplicated because of float precision.
    
  d[[run_i + 1]] = this_d
}
```

# PCA plots by run


```{r}
runs_to_plot = 4  # we plot a subset of the runs to make laying out the plots easier, results look similar across all
overall_plots = replicate(runs_to_plot, ggplot())
constant_poly_plots = replicate(runs_to_plot, ggplot())
single_var_poly_plots = replicate(runs_to_plot, ggplot())
meta_mapping_plots = replicate(runs_to_plot, ggplot())
meta_mapping_plots_34 = replicate(runs_to_plot, ggplot())
perm_mapping_plots = replicate(runs_to_plot, ggplot())
for (run_i in 1:runs_to_plot) {
  this_d = d[[run_i]]
  pcs = prcomp(this_d %>%
                 select(-dimension, -run), center=T, scale=T)
  pc_d = data.frame(pcs$rotation)
  pc_d$task = row.names(pc_d)
  pc_d = pc_d %>% 
    separate(task, c("task", "base_or_meta"), sep=":") %>%
    mutate(base_or_meta = gsub("_[0-9]+", "", base_or_meta), # some tasks are duplicated because of float precision.
           meta_task_type = str_extract(task, "mult|add|perm|square")) %>%
    rowwise() %>%
    mutate(vars_relevant=ifelse(base_or_meta == "base",
                                str_extract_all(task, "X[0-4]") %>%
                                  unlist %>% unique %>% sort %>%
                                  paste(collapse=""),
                                NA)) %>%
    ungroup() %>%
    mutate(num_vars_relevant=str_length(vars_relevant)/2,
           base_or_meta = case_when(base_or_meta == "base" ~ "Base (polynomial)",
                                    base_or_meta == "meta_class" ~ "Meta Class.",
                                    base_or_meta == "meta_map" ~ "Meta Mapping",
                                    T ~ base_or_meta
           ))
  
  # overall: base + meta
  overall_plots[[run_i]] = ggplot(pc_d, aes(x=PC1, y=PC2, color=num_vars_relevant, shape=base_or_meta, size=base_or_meta)) +
    geom_point() +
    #scale_color_brewer(palette="Dark2") + 
    labs(x="PC 1", y="PC 2") +
#    scale_color_brewer(palette="Purples") +
    guides(color=guide_legend(title="Number of\nrelevant variables"),
           shape=guide_legend(title="Type"),
           size=guide_legend(title="Type"))
  
  # constant polynomials organized by value
  constant_poly_plots[[run_i]] = ggplot(pc_d %>%
           filter(num_vars_relevant == 0) %>%
           mutate(constant_val = as.numeric(task),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1, y=PC2, color=compressed_constant_val)) +
    geom_point() +  
    scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-6, -4, -2, 0, 2, 4, 6),
                        labels=c(-36, -16, -4, 0, 4, 16, 36)) +
    #scale_color_gradient2() +#breaks=c(-4, -2, 0, 2, 4, 6),
                          #labels=c(-16, -4, 0, 4, 16, 36)) + 
    labs(x="PC 1", y="PC 2") +
    guides(color=guide_colorbar(title="Value\n(sqrt)"))
  
  # single variable polynomials
  single_var_poly_plots[[run_i]] = ggplot(pc_d %>%
           filter(num_vars_relevant == 1), 
         aes(x=PC1, y=PC2, color=vars_relevant)) +
    geom_point() +
    scale_color_brewer(palette="Set2") +
    labs(x="PC 1", y="PC 2") +
    guides(color=guide_legend(title="Variable"))
  
  # meta-mapping only PCA
  pcs = prcomp(this_d %>%
                 select(contains("meta_map")), center=T, scale=T)
  pc_d = data.frame(pcs$rotation)
  pc_d$task = row.names(pc_d)
  pc_d = pc_d %>% 
    separate(task, c("task", "base_or_meta"), sep=":") %>%
    mutate(meta_task_type = str_extract(task, "mult|add|perm|square"),
           task = case_when(meta_task_type %in% c("mult", "add") ~ str_extract(task, "[a-z0-9_-]+"),
                            meta_task_type == "perm" ~ paste("perm", str_extract(task, "[0-3]+")),
                            T ~ task)) %>%
    mutate(base_or_meta = case_when(base_or_meta == "base" ~ "Base (polynomial)",
                                    base_or_meta == "meta_class" ~ "Meta Class.",
                                    base_or_meta == "meta_map" ~ "Meta Mapping",
                                    T ~ base_or_meta
           ))
  
  meta_mapping_plots[[run_i]] = ggplot(pc_d, aes(x=PC1, y=PC2, color=meta_task_type)) +
#    geom_point(size=1.5) +
    geom_text(aes(label=task),
              position=position_nudge(x=0.05)) +
    scale_color_brewer(palette="Set1") + 
    labs(x="PC 1", y="PC 2") +
    guides(color=F)
  
  meta_mapping_plots_34[[run_i]] = ggplot(pc_d, aes(x=PC3, y=PC4, color=meta_task_type)) +
#    geom_point(size=1.5) +
    geom_text(aes(label=task),
              position=position_nudge(x=0.05)) +
    scale_color_brewer(palette="Set1") + 
    labs(x="PC 3", y="PC 4") +
    guides(color=F)

  # perm only PCA
  pcs = prcomp(this_d %>%
                 select(contains("perm")), center=T, scale=T)
  pc_d = data.frame(pcs$rotation)
  pc_d$task = row.names(pc_d)
  pc_d = pc_d %>% 
    separate(task, c("task", "base_or_meta"), sep=":") %>%
    mutate(meta_task_type = str_extract(task, "mult|add|perm|square"),
           task = case_when(meta_task_type %in% c("mult", "add") ~ str_extract(task, "[a-z0-9_-]+"),
                            meta_task_type == "perm" ~ paste("perm", str_extract(task, "[0-3]+")),
                            T ~ task)) %>%
    mutate(base_or_meta = case_when(base_or_meta == "base" ~ "Base (polynomial)",
                                    base_or_meta == "meta_class" ~ "Meta Class.",
                                    base_or_meta == "meta_map" ~ "Meta Mapping",
                                    T ~ base_or_meta
           ))
  
  perm_mapping_plots[[run_i]] = ggplot(pc_d, aes(x=PC1, y=PC2, color=meta_task_type)) +
#    geom_point(size=1.5) +
    geom_text(aes(label=task),
              position=position_nudge(x=0.05)) +
    scale_color_brewer(palette="Set1", drop=F) + 
    labs(x="PC 1", y="PC 2") +
    guides(color=F)
  
}
```


```{r}
ggarrange(overall_plots[[1]], overall_plots[[2]], overall_plots[[3]], overall_plots[[4]],  # plotlist arg isn't working and I don't want to debug it
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.75,
          common.legend=T)

#ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/overall_PCA.png", width=8, height=8)
ggsave("../metamapping_paper/figures/polynomial_reps/overall_PCA.pdf", width=8, height=8)
```

```{r}
ggarrange(constant_poly_plots[[1]], constant_poly_plots[[2]], constant_poly_plots[[3]], constant_poly_plots[[4]], 
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.75)

#ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/constant_poly_PCA.png", width=8, height=6)
#ggsave("../metamapping_paper/figures/polynomial_reps/constant_poly_PCA.png", width=8, height=6)
ggsave("../metamapping_paper/figures/polynomial_reps/constant_poly_PCA.pdf", width=8, height=6)
```

```{r}
ggarrange(single_var_poly_plots[[1]], single_var_poly_plots[[2]], single_var_poly_plots[[3]], single_var_poly_plots[[4]], 
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.425,
          common.legend=T)

ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/single_var_poly_PCA.png", width=8, height=8)
```

```{r}
ggarrange(meta_mapping_plots[[1]], meta_mapping_plots[[2]], meta_mapping_plots[[3]], meta_mapping_plots[[4]], 
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.425,
          common.legend=T)

#ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/meta_mapping_PCA.png", width=8, height=8)
ggsave("../metamapping_paper/figures/polynomial_reps/meta_mapping_PCA.pdf", width=8, height=8)

```

```{r}
ggarrange(meta_mapping_plots_34[[1]], meta_mapping_plots_34[[2]], meta_mapping_plots_34[[3]], meta_mapping_plots_34[[4]], 
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.425,
          common.legend=T)

ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/meta_mapping_PCA_34.png", width=8, height=8)
#ggsave("../metamapping_paper/figures/polynomial_reps/meta_mapping_PCA_34.png", width=8, height=8)

```


```{r}
ggarrange(perm_mapping_plots[[1]], perm_mapping_plots[[2]], perm_mapping_plots[[3]], perm_mapping_plots[[4]], 
          labels=sprintf("Run %i", 0:(runs_to_plot-1)),
          label.x=0.425,
          common.legend=T)

ggsave("../../psych/dissertation/2-HoMM/figures/polynomial_reps/perm_mapping_PCA.png", width=8, height=8)
#ggsave("../metamapping_paper/figures/polynomial_reps/meta_mapping_PCA.png", width=8, height=8)

```


# analyzing homoiconicity's effect on representations

```{r}
similarities = replicate(num_runs, data.frame())
control_similarities = replicate(num_runs, data.frame())
for (run_i in 1:num_runs) {
  this_d = d[[run_i]] %>%
    select(-dimension, -run) %>%
    mutate_all(function(x) {return(x / sqrt(sum(x^2)))}) %>% # norm columns
    as.matrix
  
  similarities_matrix = t(this_d) %*% this_d
  
  similarities[[run_i]] = similarities_matrix %>%
    as.data.frame %>%
    mutate(task1 = rownames(similarities_matrix)) %>%
    pivot_longer(-task1, names_to="task2", values_to="similarity") %>%
    mutate(run=run_i)
  
  control_d = this_d 
  # shuffle
  for (i in 1:ncol(this_d)) {
    control_d[, i] = sample(control_d[, i], nrow(this_d), replace=F)
  }
  for (j in 1:nrow(this_d)) {
    control_d[j, ] = sample(control_d[j, ], ncol(this_d), replace=F)
  }
  control_similarities_matrix = t(control_d) %*% control_d
  
  similarities[[run_i]] = bind_rows(
    similarities_matrix %>%
      as.data.frame %>%
      mutate(task1 = rownames(similarities_matrix)) %>%
      pivot_longer(-task1, names_to="task2", values_to="similarity") %>%
      mutate(condition = "Real similarities"),
    control_similarities_matrix %>%
      as.data.frame %>%
      mutate(task1 = rownames(similarities_matrix)) %>%
      pivot_longer(-task1, names_to="task2", values_to="similarity") %>%
      mutate(condition = "Scrambled control")
    )%>%
      mutate(run=run_i)
  
}
similarities = bind_rows(similarities)
```

## what is similarity between meta-mapping reps and basic task reps?

```{r}
ggplot(data=similarities %>%
         filter(task1 != task2) %>%
         mutate(task1_is_meta = grepl("meta", task1),
                task2_is_meta = grepl("meta", task2),
                task_grouping = case_when(!task1_is_meta & ! task2_is_meta ~ "Base vs. base",
                                          !task1_is_meta &  task2_is_meta ~ "Base vs. meta",
                                          task1_is_meta &  task2_is_meta ~ "Meta vs. meta",
                                          T ~ "NA")) %>%
         filter(task_grouping != "NA", # drop duplicates
                condition == "Real similarities" | task_grouping == "Base vs. meta") %>%
         mutate(task_grouping = ifelse(condition == "Scrambled control",
                                       "Scrambled control", task_grouping),
                task_grouping = factor(task_grouping, levels=c("Scrambled control", "Base vs. base", "Meta vs. meta", "Base vs. meta")),
                run = sprintf("Run %i", run)), 
       
       aes(x=similarity, color=task_grouping)) +
  facet_wrap(. ~ run) +
  scale_color_manual(values=c("#000000", "#b2182b", "#dfc28d", "#4eb3d3")) +
  geom_density()  +
  guides(color=guide_legend(title="Task pairing")) +
  labs(x="Cosine between representations", y="Density") +
  theme(legend.position=c(0.85, 0.25))
  
ggsave("../metamapping_paper/figures/polynomial_reps/meta_base_rep_similarity.pdf", width=8, height=4)
```

```{r}
similarities %>%
  filter(task1 != task2) %>%
  mutate(task1_is_meta = grepl("meta", task1),
         task2_is_meta = grepl("meta", task2)) %>%
  filter(!task1_is_meta & task2_is_meta) %>%
  group_by(condition, run) %>%
  summarize(mean_similarity = mean(similarity),
            mean_abs_similarity = mean(abs(similarity)))
```


## now restrict to cross-similarities
```{r}
similarities = similarities %>%
  filter(condition == "Real similarities",
         grepl("meta", task1),
         grepl("base", task2))
```


## do add/mult align with constants?

```{r}
add_mult_constant_d = similarities %>%
  filter(grepl("add_|mult_", task1),
         !grepl("X", task2)) %>%
  mutate(meta_constant_val=as.numeric(str_extract(task1, "-?[0-9]+\\.[0-9]*")),
         base_constant_val=as.numeric(str_extract(task2, "-?[0-9]+\\.[0-9]*")),
         val_product=meta_constant_val * base_constant_val,
         same_sign=sign(meta_constant_val) == sign(base_constant_val),
         meta_task_type=str_extract(task1, "add|mult"))
```

```{r}
ggplot(add_mult_constant_d,
       aes(x=base_constant_val, y=similarity, fill=meta_task_type)) +
  geom_point() +
  facet_grid(meta_task_type~ meta_constant_val)
```

```{r}
ggplot(add_mult_constant_d,
       aes(x=same_sign, y=similarity, fill=meta_task_type)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~run)
```

```{r}
add_mult_constant_lms = add_mult_constant_d %>%
  group_by(meta_task_type, run) %>%
  do(result=summary(lm(.$similarity ~ .$same_sign)))
add_mult_constant_lms
```

```{r}
add_mult_constant_lms %>% 
  filter(meta_task_type == "add") %>%
  pull(result)
```


```{r}
add_mult_constant_lms %>% 
  filter(meta_task_type == "mult") %>%
  pull(result)
```

```{r}
ggplot(add_mult_constant_d,
       aes(x=run, y=similarity, fill=same_sign)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~meta_task_type)
```

```{r}
ggplot(add_mult_constant_d %>%
         filter(meta_task_type == "mult") %>%
         mutate(same_sign = ifelse(same_sign, "Same sign", "Different sign")),
       aes(x=run, y=similarity, fill=same_sign)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  scale_fill_manual(values=c("#d53e4f", "#abdda4")) +
  guides(fill=guide_legend(title="Base + meta\nsigns")) +
  labs(y="Cosine between representations", x="Run")
ggsave("../metamapping_paper/figures/polynomial_reps/mult_const_alignment.pdf", width=4, height=3)
```


## does squaring relate to square terms?

```{r}
squaring_d = similarities %>%
  filter(grepl("square", task1)) %>%
  mutate(base_contains_square = grepl("^2", task2),
         base_degree=case_when(!grepl("X", task2) ~ 0,
                               grepl("^2|X0X|X1X|X2X|X3X", task2) ~ 2,
                               T ~ 1))
```

```{r}
ggplot(squaring_d %>%
         filter(grepl("X", task2)),
       aes(x=base_contains_square, fill=base_contains_square, y=similarity)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~run)
```


```{r}
ggplot(squaring_d,
       aes(x=base_degree, y=similarity)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~run)
```

## do variables affected by permutations relate to variables present in a polynomial?

```{r}
perm_poly_d = similarities %>%
  filter(grepl("permute", task1),
         grepl("X", task2)) %>%
  mutate(meta_permutation=str_extract(task1, "[0-3]+"),
         meta_X0=substr(meta_permutation, 1, 1),
         meta_X1=substr(meta_permutation, 2, 2),
         meta_X2=substr(meta_permutation, 3, 3),
         meta_X3=substr(meta_permutation, 4, 4),
         meta_X0_affected=meta_X0 != 0,
         meta_X1_affected=meta_X1 != 1,
         meta_X2_affected=meta_X2 != 2,
         meta_X3_affected=meta_X3 != 3,
         base_X0_relevant = grepl("X0", task2),
         base_X1_relevant = grepl("X1", task2),
         base_X2_relevant = grepl("X2", task2),
         base_X3_relevant = grepl("X3", task2),
         meta_perm_is_transposition = meta_permutation %in% c("1023", "0213", "0132", "2103", "3120", "0321") # switches only two variables

  )
```

```{r}
ggplot(perm_poly_d %>%
         filter(meta_perm_is_transposition) %>%
         mutate(         
           transposed_variables_relevant = case_when(meta_permutation == "1023" & base_X0_relevant & base_X1_relevant ~ T, 
                                                     meta_permutation == "0213" & base_X1_relevant & base_X2_relevant ~ T, 
                                                     meta_permutation == "0132" & base_X2_relevant & base_X3_relevant ~ T, 
                                                     meta_permutation == "2103" & base_X0_relevant & base_X2_relevant ~ T, 
                                                     meta_permutation == "3120" & base_X0_relevant & base_X3_relevant ~ T, 
                                                     meta_permutation == "0321" & base_X1_relevant & base_X3_relevant ~ T, 
                                                    T ~ F)),
       aes(x=transposed_variables_relevant, fill=transposed_variables_relevant, y=similarity)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~run)
```

```{r}
ggplot(perm_poly_d %>%
         filter(meta_perm_is_transposition) %>%
         mutate(         
           preserved_variables_relevant = case_when(meta_permutation == "1023" & base_X2_relevant & base_X3_relevant ~ T, 
                                                    meta_permutation == "0213" & base_X0_relevant & base_X3_relevant ~ T, 
                                                    meta_permutation == "0132" & base_X0_relevant & base_X1_relevant ~ T, 
                                                    meta_permutation == "2103" & base_X1_relevant & base_X3_relevant ~ T, 
                                                    meta_permutation == "3120" & base_X1_relevant & base_X2_relevant ~ T, 
                                                    meta_permutation == "0321" & base_X0_relevant & base_X2_relevant ~ T, 
                                                    T ~ F)),
       aes(x=preserved_variables_relevant, fill=preserved_variables_relevant, y=similarity)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_linerange(stat="summary",
                 fun.data="mean_cl_boot",
                 position=position_dodge(0.9)) +
  facet_wrap(~run)
```


# looking at PCA subspace similarities

```{r}
all_simil_data = replicate(runs_to_plot, data.frame())
for (run_i in 1:runs_to_plot) {
  this_d = d[[run_i]]
  centered_d = this_d %>% 
    pivot_longer(-c(dimension, run), names_to="task", values_to="value") %>%
    group_by(dimension, run) %>%
    mutate(value = value - mean(value)) %>%
    ungroup() %>%
    pivot_wider(names_from=task, values_from=value)
    
  pcs = prcomp(centered_d %>%
                 select(-dimension, -run, -contains("meta")) %>% 
                 t(), center=F, scale=F)
  this_base_sds = pcs$sdev
  this_base_pcs = pcs$rotation

  
  # meta-mapping only PCA
  pcs = prcomp(centered_d %>%
                 select(contains("meta_map")) %>%
                 t(), center=F, scale=F)
  this_meta_sds = pcs$sdev
  this_meta_pcs = pcs$rotation
  
  PC_simils = t(this_base_pcs) %*% this_meta_pcs
  colnames(PC_simils) = paste("meta_", colnames(PC_simils), sep="")
  rownames(PC_simils) = paste("base_", rownames(PC_simils), sep="")
  
  # control
  control_d = centered_d %>%
                select(contains("meta_map")) %>%
                t()
  for (i in 1:512) {
    control_d[, i] = sample(control_d[, i], 20, replace=F)
  }
  for (j in 1:20) {
    control_d[j, ] = sample(control_d[j, ], 512, replace=F)
  }
  control_pcs = prcomp(control_d,
                       center=F, scale=F)
  control_meta_sds = control_pcs$sdev
  control_meta_pcs = control_pcs$rotation
  
  control_PC_simils = t(this_base_pcs) %*% control_meta_pcs
  colnames(control_PC_simils) = paste("meta_", colnames(control_PC_simils), sep="")
  rownames(control_PC_simils) = paste("base_", rownames(control_PC_simils), sep="")
  
  this_meta_variances = this_meta_sds^2
  this_meta_proportion_variances = this_meta_variances / sum(this_meta_variances)
  
  this_control_meta_variances = control_meta_sds^2
  this_control_meta_proportion_variances = this_control_meta_variances / sum(this_control_meta_variances)
  
  PC_simil_data = as.data.frame(PC_simils) %>%
    mutate(base_PC = row.names(.)) %>%
    mutate(condition = "Original")
  
  control_PC_simil_data = as.data.frame(control_PC_simils) %>%
    mutate(base_PC = row.names(.)) %>%
    mutate(condition = "Scrambled")
  
  all_simil_data[[run_i]] = bind_rows(PC_simil_data, control_PC_simil_data) %>%
    pivot_longer(starts_with("meta"), names_to="meta_PC", values_to="cosine_similarity") %>%
    mutate_at(vars(c("base_PC", "meta_PC")), function(x) {as.numeric(str_extract(x, "[0-9]+"))}) %>%
    mutate(run=run_i) %>%
    left_join(data.frame(meta_PC=c(1:length(this_meta_proportion_variances), 1:length(this_control_meta_proportion_variances)), 
                         meta_percent_variance=c(this_meta_proportion_variances, this_control_meta_proportion_variances),
                         condition=c(rep("Original", length(this_meta_proportion_variances)), rep("Scrambled", length(this_control_meta_proportion_variances)))))
}
all_simil_data = bind_rows(all_simil_data)
```

```{r}
all_simil_data = all_simil_data %>%
  mutate(plotting_run = sprintf("Run %i", run),
         plotting_condition = factor(condition, levels=c("Original", "Scrambled")))
```

## plotting comparison between original and control projections of top 20 modes 


Will color relative to the max similarity under the control and original distributions.

```{r}
cutoff = all_simil_data %>% 
  filter(condition == "Scrambled",
         base_PC <= 20) %>%
  pull(cosine_similarity) %>%
  abs %>%
  max

max_cutoff = all_simil_data %>% 
  filter(condition == "Original",
         base_PC <= 20) %>%
  pull(cosine_similarity) %>%
  abs %>%
  max
```

```{r}
ggplot(all_simil_data %>%
         filter(base_PC <= 20) %>%
         mutate(condition=factor(condition, levels=c("Original", "Scrambled"),
                                 labels=c("Original", "Scrambled control"))),
       aes(x=base_PC, y=meta_PC, fill=abs(cosine_similarity))) +
  geom_raster() + 
  scale_fill_gradient2(low="#FFFFFF", mid="#FFDD88", high="#008800", midpoint=cutoff,
                       limits=c(0., NA),
                       breaks=c(0, cutoff, max_cutoff), 
                       labels=c("0.", sprintf("%.2f - max.\nin scrambled", cutoff), sprintf("%.2f - max.\nin original", max_cutoff))) +
  facet_grid(condition ~ plotting_run) +
  labs(x="Base P.C. index", y="Meta P.C. index") +
  guides(fill=guide_colorbar(title="Alignment\n(Abs. cosine)\n"))

ggsave("../metamapping_paper/figures/polynomial_reps/top_PC_alignments.pdf", width=8, height=4)
```

## how much variance of meta reps is contained in first few principle components of the base reps?

```{r}
shared_variance_d = all_simil_data %>%
  mutate(variance_contained = abs(cosine_similarity)^2 * meta_percent_variance) %>% 
  group_by(plotting_run, condition, base_PC) %>%
  summarize(PC_variance_contained = sum(variance_contained)) %>%
  group_by(plotting_run, condition) %>%
  mutate(cumulative_variance_contained = cumsum(PC_variance_contained)) %>%
  ungroup()
```

```{r}
ggplot(data=shared_variance_d %>%
         mutate(condition=factor(condition, levels=c("Original", "Scrambled"),
                                 labels=c("Original", "Scrambled\ncontrol"))),
       aes(x=base_PC, y=cumulative_variance_contained, color=condition)) +
  geom_line() +
  facet_wrap(~plotting_run) +
  scale_color_manual(values=c("#005500", "#EECC77")) +
  labs(x="Base PC index",
       y="Cumulative meta representation variance explained")
  
ggsave("../metamapping_paper/figures/polynomial_reps/cumulative_PC_variance.pdf", width=8, height=4)
```


# mapped representations plots


```{r message=FALSE, warning=FALSE}
mapped_results_dir = "./polynomial_mapped_rep_results"
mappings_to_load = c("add_3.000000", "mult_3.000000", "permute_1302", "permute_3210", "square")
num_mapped_runs = 4
mapped_d = replicate(num_mapped_runs, data.frame())
meta_pairing_d = replicate(num_mapped_runs, data.frame())
for (run_i in 0:(num_mapped_runs-1)) {
  this_d = read_csv(sprintf("%s/run%i_task_representations.csv", mapped_results_dir, run_i)) 
  this_d = this_d %>%
    mutate(run = run_i)# %>%
  curr_colnames = colnames(this_d)
  colnames(this_d) = c("dimension", paste(curr_colnames[2:(length(curr_colnames) - 1)], sprintf(":%s", "None"), sep=""), "run")
  for (this_mapping in mappings_to_load) {
    this_mapped_d = read_csv(sprintf("%s/run%i_task_representations__mapped-%s.csv", mapped_results_dir, run_i, this_mapping)) 
    this_mapped_d = this_mapped_d %>%
      select(-dimension)
    colnames(this_mapped_d) = paste(colnames(this_mapped_d), sprintf("_train:mapping_%s", this_mapping), sep="")
    this_d = bind_cols(this_d, 
                       this_mapped_d)
  }
  mapped_d[[run_i + 1]] = this_d
  
  # pairings
  this_pairing_d = read_csv(sprintf("%s/run%i_meta_pairings.csv", mapped_results_dir, run_i)) %>%
    mutate(run = run_i)
  meta_pairing_d[[run_i + 1]] = this_pairing_d
    
}
```

## transformation plots


```{r}
run_i = 4
this_d = mapped_d[[run_i]]
pcs = prcomp(this_d %>%
               select(-dimension, -run), center=T, scale=T)
pc_d = data.frame(pcs$rotation)
pc_d$task = row.names(pc_d)
pc_d = pc_d %>% 
  separate(task, c("task", "base_or_meta", "mapping"), sep=":") %>%
  filter(!grepl("[0-9]", base_or_meta)) %>% # drop duplicates
  separate(base_or_meta, c("base_or_meta", "base_task_toe"), sep="_") %>%
  rowwise() %>%
  mutate(vars_relevant=ifelse(base_or_meta == "base",
                              str_extract_all(task, "X[0-4]") %>%
                                unlist %>% unique %>% sort %>%
                                paste(collapse=""),
                              NA)) %>%
  ungroup() %>%
  mutate(num_vars_relevant=str_length(vars_relevant)/2,
         base_or_meta = case_when(base_or_meta == "base" ~ "Base (polynomial)",
                                  base_or_meta == "meta_class" ~ "Meta Class.",
                                  base_or_meta == "meta_map" ~ "Meta Mapping",
                                  T ~ base_or_meta
         )) %>%
  pivot_longer(starts_with("PC"), names_to="PC") %>%
  unite(PC_mapping, PC, mapping) %>%
  pivot_wider(names_from=PC_mapping, values_from=value)



```

```{r}
ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train") %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_add_3.000000`, yend=`PC2_mapping_add_3.000000`, color=compressed_constant_val)) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.1, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=F)#guide_colorbar(title="Polynomial\nconstant\nterm")) +
  #theme(legend.position=c(0.9, 0.75))
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_add_3.pdf", width=4, height=4)
```

```{r}
  ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train") %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_mult_3.000000`, yend=`PC2_mapping_mult_3.000000`, color=compressed_constant_val, shape=as.factor(num_vars_relevant))) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.1, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=guide_colorbar(title="Polynomial\nconstant\nterm")) +
  theme(legend.position=c(0.9, 0.75))
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_mult_3.pdf", width=4, height=4)
```

```{r}
p = ggplot(pc_d %>%
         filter(base_or_meta == "Base (polynomial)",
                base_task_toe == "train") %>%
         mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                squareable = !grepl("^2|X[0-3]X", task),
                compressed_constant_val = ifelse(constant_val > 0,
                                                 sqrt(constant_val),
                                                 -sqrt(-constant_val)),
                is_constant = ifelse(num_vars_relevant == 0, "Constant polynomials", "Non-constant polynomials")) %>%
           filter(squareable), 
       aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_square`, yend=`PC2_mapping_square`, color=compressed_constant_val, shape=as.factor(num_vars_relevant))) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.1, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=guide_colorbar(title="Polynomial\nconstant\nterm"))  +
  theme(legend.position=c(0.9, 0.75))#+
#  facet_wrap(~is_constant)
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_square.pdf", plot=p, width=4, height=4)
#ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_square_faceted.png", 
#       plot=p + facet_wrap(~is_constant), width=10, height=6)
p
```

```{r}
ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train",
                  num_vars_relevant > 0) %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_permute_3210`, yend=`PC2_mapping_permute_3210`, color=compressed_constant_val, shape=as.factor(num_vars_relevant))) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.1, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=guide_colorbar(title="Polynomial\nconstant\nterm")) +
  theme(legend.position=c(0.9, 0.75)) #+
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_permute_3210.pdf", width=4, height=4)
```

```{r}
ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train",
                  num_vars_relevant > 0) %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_permute_1302`, yend=`PC2_mapping_permute_1302`, color=compressed_constant_val, shape=as.factor(num_vars_relevant))) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.1, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=guide_colorbar(title="Polynomial\nconstant\nterm")) +
  theme(legend.position=c(0.9, 0.75)) #+
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_permute_1302.pdf", width=4, height=4)
```

## compare mapped to target

```{r}
original_rep_d = pc_d %>% 
  select(task, PC1_None, PC2_None) %>%
  group_by(task) %>% # remove redundancies
  summarize(PC1_None = head(PC1_None, 1),  
            PC2_None = head(PC2_None, 1))  %>% 
  ungroup()
```


```{r}
joined_d = pc_d %>%
  filter(base_task_toe == "train") %>%  # only train tasks get mapped, eval will be joined as targets
  select(-base_task_toe) %>%
  left_join(meta_pairing_d[[run_i]] %>%
              filter(meta_mapping %in% mappings_to_load) %>%
              select(-meta_toe) %>%  # only trained, for now
              mutate(meta_mapping = paste("mapping_", meta_mapping, "_target", sep="")) %>%
              unite(input_base_toe, input, base_toe, sep=":") %>%
              spread(meta_mapping, target) %>%
              separate(input_base_toe, c("task", "base_toe"), sep=":")) %>%
  # now get PCs for true targets
  left_join(original_rep_d %>%
              rename(PC1_square_target=PC1_None,
                     PC2_square_target=PC2_None),
            by=c("mapping_square_target"="task")) %>%
  left_join(original_rep_d %>%
              rename(PC1_mult_3_target=PC1_None,
                     PC2_mult_3_target=PC2_None),
            by=c("mapping_mult_3.000000_target"="task")) %>%
  left_join(original_rep_d %>%
              rename(PC1_add_3_target=PC1_None,
                     PC2_add_3_target=PC2_None),
            by=c("mapping_add_3.000000_target"="task")) %>%
  left_join(original_rep_d %>%
              rename(PC1_permute_1302_target=PC1_None,
                     PC2_permute_1302_target=PC2_None),
            by=c("mapping_permute_1302_target"="task"))
  
```

For now, simplify for plotting
```{r}
plotting_joined_d = joined_d %>%
  filter(base_or_meta == "Base (polynomial)") %>%#,
        # (!is.na(PC1_square_target)) | (!is.na(PC1_mult_3_target)) | (!is.na(PC1_add_3_target)) | (!is.na(PC1_permute_1302_target))) %>%
  mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++(?!X)")),
         constant_val = ifelse(is.na(constant_val), 0, constant_val),
         squareable = !grepl("^2|X[0-3]X", task),
         compressed_constant_val = ifelse(constant_val > 0,
                                          sqrt(constant_val),
                                          -sqrt(-constant_val))) %>%
  select(task, base_or_meta, base_toe, vars_relevant, num_vars_relevant, compressed_constant_val, squareable, starts_with("PC1_"), starts_with("PC2_")) %>%
  rename(source_PC1=PC1_None,
         source_PC2=PC2_None) %>%
  pivot_longer(starts_with("PC"), names_to="PC_and_mapping") %>%
  mutate(mapping=str_extract(PC_and_mapping, "add_3|mult_3|square|permute_1302|permute_3210|None"),
         PC=str_extract(PC_and_mapping, "PC1|PC2"),
         type=ifelse(grepl("target", PC_and_mapping), "target_output", "actual_output")) %>%
  select(-PC_and_mapping) %>%
  pivot_wider(names_from=PC, values_from=value)
         
```

TODO: think about whether it's more appropriate to compare on evals (where mapping hasn't been trained, but also ground-truth target isn't known, because target task isn't trained) or on trains (where mapping is trained but at least we know target performance is good).

```{r}
# legend drawing function with arrow fill bugfix, copied from https://stackoverflow.com/questions/57193353/show-filled-arrow-in-legend-ggplot
library(grid)
library(rlang)

draw_key_segment_custom <- function(data, params, size) {
  if (is.null(data$linetype)) {
    data$linetype <- 0
  } else {
    data$linetype[is.na(data$linetype)] <- 0
  }

  segmentsGrob(0.1, 0.5, 0.9, 0.5,
    gp = gpar(
      col = alpha(data$colour %||% data$fill %||% "black", data$alpha),
      # the following line was added relative to the ggplot2 code
      fill = alpha(data$colour %||% data$fill %||% "black", data$alpha),
      lwd = (data$size %||% 0.5) * .pt,
      lty = data$linetype %||% 1,
      lineend = "butt"
    ),
    arrow = params$arrow
  )
}
```



```{r}
square_match_plotting_d = plotting_joined_d %>%
  filter(mapping == "square",
         squareable,
         base_toe == "train",
         task %in% c("-3.55", "3.54X1", "1.56+1.46X2", "-2.18+1.10X1", "3.97", "-1.60+-1.64X3")) %>%
  mutate(type = factor(type, levels=c("target_output", "actual_output"), labels=c("Nominal\ntarget", "Actual\noutput")),
         task = task %>%
           gsub("X0", "w", .) %>%
           gsub("X1", "x", .) %>%
           gsub("X2", "y", .) %>%
           gsub("X3", "z", .) %>%
           gsub("\\+-", "-", .))


square_match_plot = ggplot(square_match_plotting_d, 
       aes(x=source_PC1, y=source_PC2, xend=PC1, yend=PC2, color=compressed_constant_val, linetype=type, alpha=type)) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.06, "inches"), type="closed")) +
  geom_text(data=square_match_plotting_d %>% 
              filter(type == "Actual\noutput") %>%
              mutate(hjust=c(0.85, 0.45, 1, 0.9, 0.5, 0.5),
                     vjust=c(0.66, 0.5, 0.5, 0.5, 0.75, 1)),
            aes(label=task,
                hjust=hjust,
                vjust=vjust),
            nudge_y=-0.0005,
            size=2) +
  labs(x="PC 1", y="PC 2") +
  xlim(-0.015, 0.005) +
  ylim(-0.017, 0.01)+
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-5, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  scale_linetype_manual(values=c("11","solid")) +
  scale_alpha_manual(values=c(0.5, 1)) +
  guides(color=F,#guide_colorbar(title="Polynomial\nconstant\nterm"),
         alpha=F,#guide_legend(title=NULL),
         linetype=F)+#guide_legend(title=NULL)) +
  theme(#legend.key.size = unit(1, "cm"),
        #legend.position=c(0.9, 0.7),
        axis.text.x=element_blank(),  # not really meaningful for principle components
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA))
square_match_plot + 
  guides(color=guide_colorbar(title="Polynomial\nconstant\nterm"),
         alpha=guide_legend(title="Mapping"),
         linetype=guide_legend(title="Mapping"))
#ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_match_square.png", width=4.33, height=3.66)
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_match_square.pdf", width=4.33, height=3.66)
```

```{r}
  ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train") %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  squareable = !grepl("^2|X[0-3]X", task),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))) %>%
           filter(squareable) %>%
           mutate(type = factor("actual_output", levels=c("target_output", "actual_output"), labels=c("Nom.\ntarget", "Actual\noutput"))), 
       aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_square`, yend=`PC2_mapping_square`, color=compressed_constant_val, linetype=type, alpha=type)) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.06, "inches"), type="closed"),
               key_glyph=draw_key_segment_custom) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  scale_linetype_manual(values=c("11","solid"), drop=F) +
  scale_alpha_manual(values=c(0.5, 1), drop=F) +
  guides(color=guide_colorbar(title="Poly.\nconstant\ncoeff."),
         linetype=guide_legend(title="Mapping"),
         alpha=guide_legend(title="Mapping")) +
  theme(axis.text.x=element_blank(),  # not really meaningful for principle components
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.margin=margin(c(5,0,5,-5)),
        legend.box.spacing=unit(0.2, "cm"),
        legend.key.size = unit(0.5, "cm")) + 
  # now insert inset plot
  annotation_custom(ggplotGrob(square_match_plot), xmin = -0.0035, xmax = 0.006, 
                    ymin = -0.002, ymax = 0.012)

ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_square_with_inset.pdf", width=4.33, height=3.66)
```

```{r}
mult_match_plotting_d = plotting_joined_d %>%
  filter(mapping == "mult_3",
         base_toe == "train",
         task %in% c("-3.55", "3.54X1", #"1.56+1.46X2", 
                     #"-2.18+1.10X1", 
                     "3.97", "-1.60+-1.64X3", "0.43+0.34X0^2", "2.39+3.07X2+0.20X1^2+-2.23X2^2")) %>%
  mutate(type = factor(type, levels=c("target_output", "actual_output"), labels=c("Nominal\ntarget", "Actual\noutput")),
         #task = ifelse(task == "2.39+3.07X2+0.20X1^2+-2.23X2^2", "2.39+3.07X2+\n0.20X1^2+-2.23X2^2", task),
         task = task %>%
           gsub("X0", "w", .) %>%
           gsub("X1", "x", .) %>%
           gsub("X2", "y", .) %>%
           gsub("X3", "z", .) %>%
           gsub("\\+-", "-", .))

mult_match_plot = ggplot(mult_match_plotting_d, 
                         aes(x=source_PC1, y=source_PC2, xend=PC1, yend=PC2, color=compressed_constant_val, linetype=type, alpha=type)) +
  # geom_segment(data=mult_match_plotting_d %>%
  #                filter(type == "Nominal\ntarget"),
  #              arrow=arrow(angle=15, length=unit(0.06, "inches"), type="open"),
  #              arrow.fill="#ffffff") +
  # geom_segment(data=mult_match_plotting_d %>%
  #                filter(type == "Actual\noutput"),
  geom_segment(arrow=arrow(angle=15, length=unit(0.06, "inches"), type="closed")) +
  geom_text(data=mult_match_plotting_d %>% 
              filter(type == "Actual\noutput") %>%
              mutate(hjust=c(0.35, 0.35, 1, 0.3, 0.75, 0.25),
                     vjust=c(0.5, 0.5, 0.5, 0.5, 0.7, 0.5)),
            aes(label=task,
                hjust=hjust,
                vjust=vjust),
            nudge_y=-0.0005,
            size=2) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-5, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  scale_linetype_manual(values=c("11","solid")) +
  scale_alpha_manual(values=c(0.5, 1)) +
  xlim(-0.015, 0.005) +
  ylim(-0.017, 0.01)+
  guides(color=F,#guide_colorbar(title="Polynomial\nconstant\nterm"),
         alpha=F,
         linetype=F)+#guide_legend(title="Meta-mapping")) +
  theme(#legend.key.size = unit(1, "cm"),
        #legend.position=c(0.7, 0.7),
        axis.text.x=element_blank(),  # not really meaningful for principle components
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
  )
mult_match_plot
ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_match_mult_3.pdf", width=3.66, height=3.66)
```

```{r}
  ggplot(pc_d %>%
           filter(base_or_meta == "Base (polynomial)",
                  base_task_toe == "train") %>%
           mutate(constant_val = as.numeric(str_extract(task, "^[-]?[0-9]+\\.[0-9]++")),
                  compressed_constant_val = ifelse(constant_val > 0,
                                                   sqrt(constant_val),
                                                   -sqrt(-constant_val))), 
         aes(x=PC1_None, y=PC2_None, xend=`PC1_mapping_mult_3.000000`, yend=`PC2_mapping_mult_3.000000`, color=compressed_constant_val)) +
  geom_segment(arrow=arrow(angle=15, length=unit(0.06, "inches"), type="closed")) +
  labs(x="PC 1", y="PC 2") +
  scale_color_gradientn(colours=c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee090", "#eedd99", "#e0f3f8", "#abd9e9", "#74add1", "#4575b4", "#313695"),
                        breaks=c(-4, -2, 0, 2, 4),
                        labels=c(-16, -4, 0, 4, 16),
                        limits=c(-4.5, 4.5)) + 
  guides(color=F) + #guide_colorbar(title="Polynomial\nconstant\nterm"))
  theme(axis.text.x=element_blank(),  # not really meaningful for principle components
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  # now insert inset plot
  annotation_custom(ggplotGrob(mult_match_plot), xmin = -0.0035, xmax = 0.006, 
                    ymin = -0.002, ymax = 0.012)

ggsave("../metamapping_paper/figures/polynomial_reps/rep_shift_mult_3_with_inset.pdf", width=3.66, height=3.66)
```


# comparing base input reps to tasks

```{r}
input_rep_d = replicate(num_mapped_runs, data.frame())
for (run_i in 0:(num_mapped_runs-1)) {
  this_d = read_csv(sprintf("%s/run%i_input_representations.csv", mapped_results_dir, run_i)) 
  this_d = this_d %>%
    mutate(run = run_i)
  input_rep_d[[run_i + 1]] = this_d
}
```

```{r}
run_i = 1

this_task_d = mapped_d[[run_i]] %>%
  pivot_longer(-c(dimension, run), names_to="task") %>%
  filter(grepl("None", task)) %>% # unmapped only
  mutate(data_type="task_rep",
         dimension=sprintf("dim%i", dimension)) %>%
  pivot_wider(names_from=dimension)

this_task_matrix = this_task_d %>%
  # normalize_rows
  pivot_longer(starts_with("dim"), names_to="dim") %>%
  group_by(task, data_type, run) %>%
  mutate(value = value / sqrt(sum(value^2))) %>%
  ungroup() %>%
  pivot_wider(names_from="dim") %>%
  select(-task, -data_type, -run) %>%
  as.matrix

this_input_d = input_rep_d[[run_i]] %>%
  mutate(data_type="input_rep") %>%
  # filter(X1 == 0,
  #        X2 == 0,
  #        X3 == 0) %>%
  unite("data_point", starts_with("X"), sep=":") %>%
  distinct()

this_input_matrix = this_input_d %>%
  # normalize_rows
  pivot_longer(starts_with("dim"), names_to="dim") %>%
  group_by(data_point, data_type, run) %>%
  mutate(value = value / sqrt(sum(value^2))) %>%
  ungroup() %>%
  pivot_wider(names_from="dim") %>%
  select(-data_point, -data_type, -run) %>%
  as.matrix
  
similarities_matrix = this_task_matrix %*% t(this_input_matrix)

colnames(similarities_matrix) = this_input_d$data_point
simil_d = similarities_matrix %>%
  as.data.frame %>%
  mutate(task = this_task_d$task) %>%
  rowwise() %>%
  mutate(task_vars_relevant=str_extract_all(task, "X[0-4]") %>%
                                unlist %>% unique %>% sort %>%
                                paste(collapse="")) %>%
  ungroup() %>%
  pivot_longer(-c(task, task_vars_relevant), names_to="data_point") %>%
  separate(data_point, c("X0", "X1", "X2", "X3"), sep=":")

```

```{r}
simil_d = simil_d %>%
  mutate(task_base_or_meta=str_extract(task, "base|meta"),
         point_num_vars_relevant = (X0 != 0) + (X1 != 0) + (X2 != 0) + (X3 != 0),
         task_num_vars_relevant = str_length(task_vars_relevant)/2,
         task_X0_relevant=grepl("X0", task),
         task_X1_relevant=grepl("X1", task),
         task_X2_relevant=grepl("X2", task),
         task_X3_relevant=grepl("X3", task),
         point_X0_relevant=X0 != 0,
         point_X1_relevant=X1 != 0,
         point_X2_relevant=X2 != 0,
         point_X3_relevant=X3 != 0,
         vars_relevant_IoU = ifelse(task_num_vars_relevant == 0 & point_num_vars_relevant ==0, NA, ((task_X0_relevant & point_X0_relevant) + (task_X1_relevant & point_X1_relevant) +(task_X2_relevant & point_X2_relevant) +(task_X3_relevant & point_X3_relevant))/ ((task_X0_relevant | point_X0_relevant) + (task_X1_relevant | point_X1_relevant) +(task_X2_relevant | point_X2_relevant) +(task_X3_relevant | point_X3_relevant))))
```

```{r}
ggplot(data=simil_d %>%
         group_by(task_num_vars_relevant, point_num_vars_relevant) %>%
         summarize(mean_abs_simil = mean(abs(value))),
       aes(x=task_num_vars_relevant, color=as.factor(point_num_vars_relevant), y=mean_abs_simil)) +
  geom_point(position="dodge")# +
  # geom_linerange(stat="summary",
  #                fun.y="mean_cl_boot",
  #                position="dodge")
```


```{r}
simil_d = simil_d %>%
  mutate(task_X0=as.numeric(str_extract(task, "[-]?[0-9]+\\.[0-9]+(?=X0+)")),
         task_X1=as.numeric(str_extract(task, "[-]?[0-9]+\\.[0-9]+(?=X1+)")),
         task_X2=as.numeric(str_extract(task, "[-]?[0-9]+\\.[0-9]+(?=X2+)")),
         task_X3=as.numeric(str_extract(task, "[-]?[0-9]+\\.[0-9]+(?=X3+)"))) %>%
  mutate_at(vars(starts_with("task_X")), function(x) {ifelse(is.na(x), 0, x)})
```

```{r}
lm(value ~ task_X0 * as.numeric(X0) + task_X1 * as.numeric(X1) + task_X2 * as.numeric(X2)  + task_X3 * as.numeric(X3),
   data=simil_d) %>%
  summary()
```

```{r}
simil_d %>% 
  mutate(abs_value=abs(value)) %>%
  filter(abs_value > 0.05) %>%
  arrange(desc(abs_value))
```


```{r}
ggplot(data=simil_d,
       aes(x=vars_relevant_IoU, color=as.factor(point_num_vars_relevant), y=abs(value))) +
  geom_point(stat="summary",
             fun.y="mean")
```

```{r}
lm(abs(value) ~ point_num_vars_relevant + task_num_vars_relevant + vars_relevant_IoU,
  data=simil_d) %>% summary()
```

```{r}
lm(value ~ mapping_scalar * (as.numeric(X0) + as.numeric(X1) + as.numeric(X2) + as.numeric(X3)),
   data=simil_d %>%
         filter(grepl("mult", task)) %>%
         mutate(mapping_scalar=as.numeric(str_extract(task, "[-]?[0-9]+\\.[0-9]+")))) %>%
  summary()
```
